---
title: "R Notebook"
output: html_notebook
---

## Download data from the web


```{r}
fileUrl <- "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.csv?accessType=DOWNLOAD"
download.file(fileUrl, destfile = "./data/cameras.csv", method = "curl")
list.files("./data")

```

## keeping track of date of download
```{r}
dateDownloaded <- date()
dateDownloaded
```

## Reading local files

```{r}
if(!file.exists("data")){
  dir.create("data")
}
filrUrl <- "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.csv?accessType=DOWNLOAD"
download.file(filrUrl, destfile = "cameras.csv", method = "curl")
dateDownloaded <- date()
```

## Reading the data

```{r}
cameraData <- read.table("./data/cameras.csv", sep=",", header = TRUE)
```
```{r}
head(cameraData)
```
## Reading excel file
```{r}
if(!file.exists("data")){
  dir.create("data")
}
fileUrl <- "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.csv?accessType=DOWNLOAD&bom=true&format=true"
download.file(fileUrl, destfile = "./data/cameras.xlsx", method = "curl")
dateDownloaded <- date()
```
```{r}
library(xlsx)
cameraData <- read.xlsx("./data/cameras.xlsx", sheetIndex=1, header=TRUE)
head(cameraData)
```
## Reading XML data

```{r}
library(XML)
fileUrl <- "http://www.w3schools.com/xml/simple.xml"
doc <- xmlTreeParse(fileUrl, useInterna=TRUE)
rootNode <- xmlRoot(doc)
xmlName(rootNode)
```

## Reading data from JSON
```{r}
library(jsonlite)
jsonData <- fromJSON("https://api.github.com/users/jtleek/repos")
names(jsonData)
```
```{r}
names(jsonData)
```
```{r}
jsonData$owner$login
```
## writing data frames to JSON
```{r}
myjson <- toJSON(iris, pretty = TRUE)
cat(myjson)
```
## Convert back to JSON
```{r}
iris2 <- fromJSON(myjson)
head(iris2)
```
## create data tables just like data frames
```{r}
library(data.table)
DF = data.frame(x=rnorm(9), y=rep(c("a","b","c"), each=3), z=rnorm(9))
head(DF, 3)
```
```{r}
DT = data.table(x=rnorm(9), y=rep(c("a", "b", "c"), each=3),z=rnorm(9))
head(DT, 3)
```
## subsetting rows
```{r}
DT[2,]
```
```{r}
DT[DT$y=="a",]
```
## subsetting rows
```{r}
DT[c(2,3)]
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
