---
title: "Getting and cleaning data Quiz1"
output: html_notebook
author: maduabuchi Okonkwo
---

## 1.Question 1

The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv

and load the data into R. The code book, describing the variable names is here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf

How many properties are worth $1,000,000 or more?
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
if(!dir.exists("data")){
  dir.create("data")
}
download.file(fileUrl, destfile = "./data/housing.csv", method = "curl")
housing <- read.csv("./data/housing.csv")


```
```{r}
head(housing)
```
```{r}
sum(housing$VAL==24 & !is.na(housing$VAL))
```
### Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate? 

```{r}
housing$FES
```
### Download the Excel spreadsheet on Natural Gas Aquisition Program here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx

Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:

```{r}
library(xlsx)
file2Url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx "
download.file(file2Url, destfile = "./data/gasdata.xlsx", method = "curl")
dat <- read.xlsx("./data/gasdata.xlsx", sheetIndex = 1, rowIndex = 18:23, colIndex = 7:15)
sum(dat$Zip*dat$Ext,na.rm=T)

```
## 4.Question 4

Read the XML data on Baltimore restaurants from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml

How many restaurants have zipcode 21231? 

```{r}
library(XML)
file3Url <- "http://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml "
doc <- xmlTreeParse(file3Url, useInternal=TRUE)
rootNode <- xmlRoot(doc)
sum(xpathSapply(rootNode, "//zipcode", xmlValue)==21231)
downloaded <- date()
downloaded
```

##The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv

using the fread() command load the data into an R object

The following are ways to calculate the average value of the variable

broken down by sex. Using the data.table package, which will deliver the fastest user time?

mean(DT$pwgtp15,by=DT$SEX)

sapply(split(DT$pwgtp15,DT$SEX),mean)

tapply(DT$pwgtp15,DT$SEX,mean)

DT[,mean(pwgtp15),by=SEX]

mean(DT[DT$SEX==1,]$pwgtp15); mean(DT[DT$SEX==2,]$pwgtp15)

rowMeans(DT)[DT$SEX==1]; rowMeans(DT)[DT$SEX==2]

```{r}
library(data.table)
file5Url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
download.file(file5Url, destfile = "./data/communitydata.csv", method = "curl")
DT <- fread("./data/communitydata.csv")
downloaded <- date()
downloaded
```
```{r}
system.time(DT[,mean(pwgtp15),by=SEX])
```
```{r}
system.time(mean(DT[DT$SEX==1,]$pwgtp15))+system.time(mean(DT[DT$SEX==2,]$pwgtp15))
```
```{r}
system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
```
```{r}
system.time(mean(DT$pwgtp15,by=DT$SEX))
```
```{r}
system.time(tapply(DT$pwgtp15,DT$SEX,mean))
```
```{r}
system.time(rowMeans(DT)[DT$SEX==1])+system.time(rowMeans(DT)[DT$SEX==2])
```

